# Data
::: callout-note
## This page contains Tabs. Please click through tabs to see additional charts/info.
:::
## Description
For this project, we will be using a subset of the **Citywide Payroll Data**.\
This dataset is available via NYC OpenData and provided by the Office of Payroll Administration (OPA).

The final subset of data for this project will be focused on 2 Agencies:\
<p style="margin-top: -15px">
-   Fire Department 🧑‍🚒 🚒
-   Police Department 👮🚓
</p>

Here are the filters applied to the original data to produce our final subset of data:
```{r}
#| echo: false
#| message: false
#| warning: false


library(tibble)
library(knitr)
library(kableExtra)

payroll_data_format <- tibble(
  column_name = c(
    "Fiscal Year", 
    "Agency Name", 
    "Work Location Borough",
    "Pay Basis"
  ),
  filter = c(
    "From 2015 to 2024 (10 years)", 
    "FIRE DEPARTMENT, POLICE DEPARTMENT",
    "BRONX, BROOKLYN, MANHATTAN, QUEENS, STATEN ISLAND",
    "per Annum, per Day, per Hour"
  )
)

names(payroll_data_format) <- c("Column Name","Filter")

kable(payroll_data_format,
      # caption = "Dataset Format", 
      align = "l") |>
  kable_styling(full_width = FALSE) |>
  column_spec(1, width = "200px")
```


<!-- First, we will examine the metadata of the dataset, how to import the data and process the subset data. Finally, an analysis of missing values to identify any potential issues that may impact our analysis. -->
<!-- , followed by the decision-making process for selecting a subset of the data and transforming the original dataset with 6,225,611 rows to our final dataset with 804,630 rows. -->

```{r}
#| echo: false

rm(list = ls())
```

```{r}
#| echo: false
#| message: false
#| warning: false

# Import Libraries
library(tibble)
library(knitr)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(visdat)
library(HH)
library(lubridate)
```

\
\



### Data information

The payroll data is provided annually by [NYC OpenData](https://data.cityofnewyork.us/City-Government/Citywide-Payroll-Data-Fiscal-Year-/k397-673e/about_data) and contains New York City employee salaries by Fiscal Year. We are using the latest snapshot of the data, released on October 30, 2024. Here are some details about the format of the data, the frequency of updates, dimensions about the data, and other relevant metadata:


::: panel-tabset
## Data Metadata

```{r}
#| echo: false


library(tibble)
library(knitr)
library(kableExtra)

# Create a tibble for the metadata
metadata <- tibble::tibble(
  Field = c(
    "Dataset Name",
    "Provided by",
    "Data Category",
    "Frequency of Updates ", 
    "Date Created", 
    "Data Last Updated", 
    "Dimensions",
    "Each row represents",
    "Source URL"
  ),
  Value = c(
    "Citywide Payroll Data (Fiscal Year)",
    "Office of Payroll Administration (OPA)",
    "City Government",
    "Annually", 
    "October 31, 2015", 
    "October 30, 2024", 
    "Raw Data: 6,225,611 rows | 17 columns <br> \n Subset Data: 804,630 rows | 17 columns",
    "City Employee Salary per Fiscal Year",
    "https://data.cityofnewyork.us/"
  )
)

# Print the metadata table using kable
kable(metadata, 
      col.names = NULL, 
      escape = F,
      format = "html",
      # caption = "Dataset Metadata",
      align = "l")  |>
  column_spec(1, width = "200px") |>
  kable_styling()

```

\
\

## Data Format

```{r}
#| echo: false



payroll_data_format <- tibble(
  column_name = c(
    "Fiscal Year", "Payroll Number", "Agency Name", "Last Name", "First Name", 
    "Mid Init", "Agency Start Date", "Work Location Borough", 
    "Title Description", "Leave Status as of June 30", "Base Salary", 
    "Pay Basis", "Regular Hours", "Regular Gross Paid", "OT Hours", 
    "Total OT Paid", "Total Other Pay"
  ),
  description = c(
    "Fiscal Year", "Payroll Number", "The Payroll agency that the employee works for", 
    "Last name of employee", "First name of employee", 
    "Middle initial of employee", "Date which employee began working for their current agency", 
    "Borough of employee's primary work location", "Civil service title description of the employee", 
    "Status of employee as of the close of the relevant fiscal year: Active, Ceased, or On Leave", 
    "Base Salary assigned to the employee", 
    "Lists whether the employee is paid on an hourly, per diem or annual basis", 
    "Number of regular hours employee worked in the fiscal year", 
    "The amount paid to the employee for base salary during the fiscal year", 
    "Overtime Hours worked by employee in the fiscal year", 
    "Total overtime pay paid to the employee in the fiscal year", 
    "Includes any compensation in addition to gross salary and overtime pay, i.e., Differentials, lump sums, uniform allowance, meal allowance, retroactive pay increases, settlement amounts, and bonus pay, if applicable."
  ),
  # api_field_name = c(
  #   "fiscal_year", "payroll_number", "agency_name", "last_name", "first_name", 
  #   "mid_init", "agency_start_date", "work_location_borough", 
  #   "title_description", "leave_status_as_of_june_30", "base_salary", 
  #   "pay_basis", "regular_hours", "regular_gross_paid", "ot_hours", 
  #   "total_ot_paid", "total_other_pay"
  # ),
  data_type = c(
    "Number", "Number", "Text", "Text", "Text", 
    "Text", "Timestamp", "Text", 
    "Text", "Text", "Number", 
    "Text", "Number", "Number", "Number", 
    "Number", "Number"
  )
)

names(payroll_data_format) <- c("Column Name","Description","Data Type")

kable(payroll_data_format,
      # caption = "Dataset Format", 
      align = "l") |>
  kable_styling(full_width = FALSE) |>
  column_spec(1, width = "200px")
```

:::

\


### Data Source


```{mermaid}
%%| echo: false

flowchart LR
  A[Importing Data] --> B[Understanding and \n cleaning the raw data]
  B --> C(Transforming columns)
  B --> D(Filtering Data)
  B --> E(Backfilling NA)
  C --> F[Saving Subset Data]
  D --> F[Saving Subset Data]
  E --> F[Saving Subset Data]
```





::: panel-tabset
## **Importing the data**\

To access the data, we can go directly to the URL:\
<p style="font-size:95%; font-style:italic">
<https://data.cityofnewyork.us/City-Government/Citywide-Payroll-Data-Fiscal-Year-/k397-673e/about_data>
</p>

From this URL, you can click the button 'Export', and then select Download File-CSV format (All data 6225611 rows). This will download a file of about 840MB. Remember, we are using the data Last Updated on October 30, 2024.

The file will be downloaded with the name: Citywide_Payroll_Data\_\_Fiscal_Year\_\_**YYYYMMDD**.csv\
<p style="font-size:90%; font-style:italic">
(YYYYMMDD refers to the date that you downloaded the file)
</p>

Once the data was downloaded, it was added to a new folder inside the repository called 'data_source'.

```{r, eval=FALSE}
#| message: false
#| warning: false

# import original data
payroll_data_source <- read.csv("./data_source/Citywide_Payroll_Data__Fiscal_Year__20241111.csv")

# make a copy of original data for transformations
payroll_data <- data.frame(payroll_data_source)

# rename columns
names(payroll_data) <- gsub("[\\.]+", "_", tolower(names(payroll_data_source)))
```


\
\

## **Cleaning the data**\

The decision-making process for selecting a subset of the data and the detailed transformation and clean up of the original dataset with 6,225,611 rows to our final dataset with 804,630 rows can be found in the `Appendix` section.


\
\

## **Creating the final subset data**\

The data used for this analysis is a subset of the entire raw data. The **final subset data** (focused on the **Fire and Police departments**) contains 804,630 rows out of the 6,225,611 rows from the original dataset. 

To generate the final subset that will be used in the analysis, please run the code below:

```{r, eval=FALSE}
# `payroll_data` dataframe comes from the **Importing the data** section.

payroll_data_fire_police <-
  payroll_data |>
  # clean agency name:
  mutate(agency_name_clean = trimws(gsub("#\\d+$", "", agency_name))) |>
  # clean job titles:
  mutate(title_description = toupper(title_description)) |> 
  # clean location name:
  mutate(work_location_borough_clean = toupper(work_location_borough)) |>

  # filtering data:
  filter(agency_name_clean %in%  c("POLICE DEPARTMENT","FIRE DEPARTMENT")) |>
  filter(work_location_borough_clean %in%   c("BRONX", "BROOKLYN", "MANHATTAN", "QUEENS", "RICHMOND")) |>
  filter(fiscal_year >= "2015" & fiscal_year <= "2024") |>
  filter(pay_basis %in%   c("per Annum", "per Day", "per Hour")) |>

  # backfill payroll number after filtering
  mutate(
    payroll_number_clean = case_when(
      is.na(payroll_number) & agency_name_clean == "FIRE DEPARTMENT" ~ 57,
      is.na(payroll_number) & agency_name_clean == "POLICE DEPARTMENT" ~ 56,
      TRUE ~ payroll_number
    )) |>

  # converting dates:
  mutate(agency_start_date_clean = mdy(agency_start_date)) |>

  # renaming clean columns to original names:
  mutate(
    agency_name = agency_name_clean,
    work_location_borough = work_location_borough_clean,
    payroll_number = payroll_number_clean,
    agency_start_date = agency_start_date_clean
  ) |>

  # dropping clean columns
  dplyr::select(-agency_name_clean,
                -work_location_borough_clean,
                -payroll_number_clean,
                -agency_start_date_clean)



# Please uncomment the CSV or the RDS method to save the subset data:

# Save the data CSV
# write.csv(payroll_data_fire_police, "data_source/payroll_data_fire_police.csv", row.names = FALSE)

# Save the data RDS
# saveRDS(payroll_data_fire_police, "data_source/payroll_data_fire_police.rds")

# Read the data RDS
# payroll_data_fire_police <- readRDS("data_source/payroll_data_fire_police.rds")
```


:::

\
\




## Missing value analysis

To analyse missing data, we will classify our data in 4 categories: \
&nbsp;&nbsp;&nbsp; [■]{style="color:#D7191C;"}  NA      - any values equal to *NA* \
&nbsp;&nbsp;&nbsp; [■]{style="color:#FDAE61;"}  Blank   - string values equal to "" \
&nbsp;&nbsp;&nbsp; [■]{style="color:#ABD9E9;"}  Zero    - numeric values equal to 0 \
&nbsp;&nbsp;&nbsp; [■]{style="color:#2C7BB6;"}  Complete - any values with data \


### NAs and Blanks

First, let's check the missing values of the original dataset.\
Using a sample of 1 million records (from 6.22 million rows) we have close to **4.5%** of missing data classified as [■]{style="color:#D7191C;"} NA or [■]{style="color:#FDAE61;"} Blank. We identified missing values in 7 columns*, including three related to a person’s name (`first_name`, `mid_init`, `last_name`), `payroll_number`, `work_location_borough`, `title_description`, and `agency_start_date`. \
<!-- <p style="margin-top: -15px;"> -->
<!-- -  3 columns related to a person's name (first_name, mid_init, last_name) -->
<!-- -  payroll_number -->
<!-- -  work_location_borough -->
<!-- -  title_description -->
<!-- -  agency_start_date \ -->
<!-- </p> -->
[**(\*based on the black highlights and the percentages listed in the column names of the vis_miss() plot)**]{style="font-size:80%;margin-top: -15px;"}



```{r}
#| echo: false
#| message: false
#| warning: false

# import original data
payroll_data_source <- read.csv("./data_source/Citywide_Payroll_Data__Fiscal_Year__20241111.csv")

# make a copy of original data for transformations
payroll_data <- data.frame(payroll_data_source)

# rename columns
names(payroll_data) <- gsub("[\\.]+", "_", tolower(names(payroll_data_source)))
```


```{r}
#| fig-width: 9
#| fig-height: 6

payroll_data_missing <- 
  payroll_data |>
  sample_n(100000) |>
  mutate(across(everything(), ~ if_else(. == "", NA, .)))

payroll_data_missing |> 
  vis_miss(
    sort_miss = TRUE,
    # cluster = TRUE, # disabled because vector memory limit of 16.0 Gb reached, see mem.maxVSize()
    show_perc_col = TRUE,
    warn_large_data = FALSE,
    ) + 
  ggtitle("Missing Values (NA or Blank) - Original Data Sample")+
  labs(caption = "Clustering has been disabled due high volume data and memory limits.") +
  theme(plot.margin = margin(r = 50))
```

\
After transforming our original data to create our final subset data (Fire and Police Departments), we can now examine the missing values of the final dataset.



```{r}
#| fig-width: 9
#| fig-height: 7
#| message: false
#| warning: false

# install.packages("remotes")
# remotes::install_github("jtr13/redav")
library(redav)

plot_missing_modified <- function(x, percent = TRUE, num_char = 100, 
                         max_rows = NULL, max_cols = NULL) {	

  na_count_all <- data.frame(is.na(x)) %>%	
    dplyr::group_by_all() %>%	
    dplyr::count(name = "count", sort = TRUE) %>%	
    dplyr::ungroup() %>%	
    tibble::rownames_to_column("pattern")	
  
  if (!is.null(max_rows)) na_count_all <- na_count_all %>% 
      dplyr::filter(as.numeric(pattern) <= max_rows)
  
  na_count_all <- na_count_all %>% 
    dplyr::mutate(pattern = factor(.data$pattern, levels = nrow(na_count_all):1))

  
  # count the number of columns with missing values; will be used later to determine if there's a "none missing" pattern	
  na_count_all <- na_count_all %>% 	
    dplyr::rowwise() %>%	
    dplyr::mutate(num_missing_cols = sum(dplyr::c_across(where(is.logical))))	
  
  # data frame for missing patterns bar chart	
  na_count_by_pattern <- na_count_all[,c("pattern", "count", "num_missing_cols")]
  na_count_by_pattern$none_missing <- ifelse(na_count_by_pattern$num_missing_cols == 0, TRUE, FALSE)

  # data frame for missing by column bar chart	
  na_count_by_column <- data.frame(is.na(x)) %>%	
    colSums() %>% 	
    sort(decreasing = TRUE) %>% 	
    tibble::enframe(name = "var", value = "count")	
  
  if (!is.null(max_cols)) na_count_by_column <- na_count_by_column %>%
    dplyr::slice(1:max_cols)
  
  # tidy and sort na_count_all by column counts	
  na_count_all_tidy <- na_count_all %>% 	
    tidyr::pivot_longer(where(is.logical), names_to = "variable") %>%	
    dplyr::filter(variable %in% na_count_by_column$var) %>%
    dplyr::mutate(variable = factor(.data$variable, levels = na_count_by_column$var))  %>% 	
    dplyr::mutate(none_missing = ifelse(.data$num_missing_cols == 0, TRUE, FALSE))
  
  # calculate the % of data shown
  perc_shown <- 100*sum(na_count_by_pattern$count)/nrow(x)
  
  # main plot
  main_plot <- na_count_all_tidy %>%
    ggplot2::ggplot(ggplot2::aes(.data$variable, .data$pattern, fill = factor(.data$value), alpha = .data$none_missing)) +	
    ggplot2::geom_tile(color = "white") +	
    ggplot2::scale_x_discrete(labels = abbreviate(levels(na_count_all_tidy$variable), num_char)) +
    ggplot2::scale_fill_manual(values = c("grey70", "mediumpurple")) +	
    ggplot2::scale_alpha_manual(values = c(.7, 1)) +	
    ggplot2::labs(y = "missing pattern",
                  caption = paste0("These missing patterns represent ",
                                   round(perc_shown, 1), "% of the rows.")) +	
    ggplot2::guides(fill = "none", alpha = "none") +	
    ggplot2::theme_classic(12)+ 
    ggplot2::theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  # check for "none missing" pattern
  none_missing_pattern <- na_count_by_pattern %>%
    dplyr::filter(.data$none_missing) %>% dplyr::pull(.data$pattern)
  
  if (length(none_missing_pattern) > 0) {	
    column_count <- min(c(ncol(na_count_all)-2, max_cols+1))
    main_plot <- main_plot +	
      ggplot2::annotate("text", x = column_count/2,	
               y = nrow(na_count_all) + 1 - as.numeric(as.character(none_missing_pattern)),	
               label = "complete cases")	
  }	
  
  # margin plots
  
  denom <- ifelse(percent, nrow(x)/100, 1)
  
  # top
  missing_by_column_plot <- 
    na_count_by_column %>% 
    dplyr::mutate(var = abbreviate(var, num_char)) %>% 
    ggplot2::ggplot(ggplot2::aes(forcats::fct_inorder(.data$var), .data$count/denom)) +	
    ggplot2::geom_col(fill = "cornflowerblue", alpha = .7) +
    ggplot2::scale_y_continuous(expand = c(0, 0), n.breaks = 3) +	
    ggplot2::labs(x = "",
                  y = ifelse(percent, "% rows \n missing:", "num rows \n missing:")) +	
    ggplot2::theme_linedraw(12) + 	
    ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),	
          panel.grid.minor.x = ggplot2::element_blank(),
          )	+ 
    ggplot2::theme(axis.text.x = element_text(angle = 45, hjust = 0))+
    scale_x_discrete(position = "top") +
    ggplot2::geom_text(
      ggplot2::aes(
        label = ifelse(.data$count / denom > 0, scales::percent(.data$count / denom / 100, accuracy = 0.1), NA)
        ),
      vjust = -0.5,
      size = 3
      ) +
    ggtitle("Missing Values (NA or Blank) - Subset Data")
  
  # right
  missing_by_pattern_plot <- 
    ggplot2::ggplot(na_count_by_pattern, ggplot2::aes(.data$pattern, .data$count/denom, alpha = .data$none_missing)) +
    ggplot2::geom_col(fill = "cornflowerblue") +
    ggplot2::coord_flip() +
    ggplot2::scale_y_continuous(expand = c(0, 0), n.breaks = 3) +
    ggplot2::scale_alpha_manual(values = c(.7, 1)) +
    ggplot2::labs(x = "", y = ifelse(percent, "% rows", "row count")) +
    ggplot2::guides(alpha = "none") +
    ggplot2::theme_linedraw(12) +
    ggplot2::theme(panel.grid.major.y = ggplot2::element_blank(), 
          panel.grid.minor.y = ggplot2::element_blank()) +
    ggplot2::geom_text(
      ggplot2::aes(
        label = ifelse(.data$count / denom > 0, scales::percent(.data$count / denom / 100, accuracy = 0.1), NA)),
      hjust = -0.1,
      size = 3
      ) 
  
  if (percent) {	
    missing_by_column_plot <- missing_by_column_plot +
      ggplot2::scale_y_continuous(expand = c(0, 0), n.breaks = 5,
                         limits = c(0, 100))	
    missing_by_pattern_plot <- missing_by_pattern_plot +
      ggplot2::scale_y_continuous(expand = c(0, 0), n.breaks = 5,
                         limits = c(0, 100))	
  }	
  
  missing_by_column_plot + patchwork::plot_spacer() + 	
    main_plot + missing_by_pattern_plot + 	
    patchwork::plot_layout(widths = c(4, 1), heights = c(2, 4))
}


payroll_data_fire_police <- readRDS("data_source/payroll_data_fire_police.rds")

payroll_data_fire_police_na_blank <-
  payroll_data_fire_police |> 
  mutate(across(everything(), ~ case_when(. == "" ~ NA, TRUE ~ .)))


plot_missing_modified(payroll_data_fire_police_na_blank, percent = TRUE)
```
::: callout-note
## plot_missing() original source function has been modified
Original function was changed to accommodate long column names in the x-axis and show % labels.
:::

\
\

### Zeros and Complete Values

In addition to NAs and Blanks, an important part of this dataset is the analysis of [■]{style="color:#ABD9E9;"} Zero values. For this dataset, Zeros are relevant information and should not be excluded from the data. Let's take a look at the full analysis of NAs, Blanks, Zeros, and Complete values. 

We can now examine how the values from the original compares to the final dataset:


::: panel-tabset
## Original VS Subset Comparison

```{r}
#| fig-width: 9
#| fig-height: 5

#Data Transformation

payroll_data_total_summary <- payroll_data |>
  summarise(across(everything(), ~ n())) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Total Count"
  )

payroll_data_na_summary <- payroll_data |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "NA"
  )

payroll_data_blank_summary <- payroll_data |>
  summarise(across(everything(), ~ sum(. == "", na.rm = TRUE))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Blank"
  )

payroll_data_zero_summary <- payroll_data |>
  summarise(across(everything(), ~ sum(. == 0, na.rm = TRUE))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Zero"
  )

payroll_data_combined_summary <- 
  payroll_data_total_summary |> 
  inner_join(payroll_data_na_summary, by = "Column") |> 
  inner_join(payroll_data_blank_summary, by = "Column") |> 
  inner_join(payroll_data_zero_summary, by = "Column")


# payroll_data_combined_summary

payroll_data_combined_summary_likert <-
payroll_data_combined_summary |> 
  mutate(`Complete` = `Total Count` - `NA` - `Blank` - Zero) |> 
  mutate(`Complete` = `Complete` / `Total Count`) |> 
  mutate(`NA` = `NA` / `Total Count`) |> 
  mutate(`Blank` = `Blank` / `Total Count`) |> 
  mutate(`Zero` = Zero / `Total Count`) |> 
  dplyr::select(`Column`,`NA`,`Blank`,Zero,`Complete`) |> 
  filter(!Column %in% c('agency_name_clean','work_location_borough_clean'))


likert_colors <- c("NA" = "#D7191C",
                   "Blank" = "#FDAE61",
                   "Zero" = "#ABD9E9",
                   "Complete" = "#2C7BB6")



payroll_data_fire_police <- readRDS("data_source/payroll_data_fire_police.rds")

payroll_data_fire_police_total_summary <- payroll_data_fire_police |>
  summarise(across(everything(), ~ n())) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Total Count"
  )

payroll_data_fire_police_na_summary <- payroll_data_fire_police |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "NA"
  )

payroll_data_fire_police_blank_summary <- payroll_data_fire_police |>
  summarise(across(everything(), ~ sum(. == "", na.rm = TRUE))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Blank"
  )

payroll_data_fire_police_zero_summary <- payroll_data_fire_police |>
  summarise(across(everything(), ~ sum(. == 0, na.rm = TRUE))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Zero"
  )


payroll_data_fire_police_combined_summary <- 
  payroll_data_fire_police_total_summary |> 
  inner_join(payroll_data_fire_police_na_summary, by = "Column") |> 
  inner_join(payroll_data_fire_police_blank_summary, by = "Column") |> 
  inner_join(payroll_data_fire_police_zero_summary, by = "Column")


payroll_data_fire_police_combined_summary_likert <-
  payroll_data_fire_police_combined_summary |> 
  mutate(`Complete` = `Total Count` - `NA` - `Blank` - Zero) |> 
  mutate(`Complete` = `Complete` / `Total Count`) |> 
  mutate(`NA` = `NA` / `Total Count`) |> 
  mutate(`Blank` = `Blank` / `Total Count`) |> 
  mutate(`Zero` = Zero / `Total Count`) |> 
  dplyr::select(`Column`,`NA`,`Blank`,Zero,`Complete`) |> 
  filter(!Column %in% c('agency_name_clean','work_location_borough_clean'))


# Chart

payroll_data_missing_group_likert <- bind_rows(
  payroll_data_combined_summary_likert |> mutate(group= 'Original Data'), 
  payroll_data_fire_police_combined_summary_likert |> mutate(group= 'Final Subset Data')
)

HH::likert(x=Column~. | group,payroll_data_missing_group_likert, 
           positive.order=TRUE,
           as.percent = T,
           main = 'Missing Value Analysis',
           xlab='Percentage',
           ylab='Columns',
           col = likert_colors,
           ReferenceZero=2.5,
           rightAxis=FALSE,
           # xlimEqualLeftRight=TRUE,
           xlim=c(-50,100),
           )
```
::: callout-note
## Diverging stacked bar chart to analyze missing data
[In this dataset, values classified as "Complete" or "Zero" are considered positive, as they indicate usable data that can be effectively analyzed. Conversely, "NA" and "Blank" values pose potential issues for data interpretation. To reflect this distinction, the diverging stacked bar chart will position "NA" and "Blank" values on the left side, representing a negative impact on the analysis.]{style="font-size:82%;"}
:::

## Original Details
```{r}
#| echo: false


payroll_data_missing_table <- tibble(
  column_name = c(
    "Payroll Number",
    "Last Name", 
    "First Name",
    "Mid Init",
    "Work Location Borough",
    "Agency Start Date",
    "Title Description",
    "Regular Hours", 
    "Regular Gross Paid", 
    "OT Hours", 
    "Total OT Paid", 
    "Total Other Pay",
    "Fiscal Year , Agency Name, Leave Status as of June 30, Base Salary, Pay Basis"
  ),
  
  
  missing_na = c(
    "28.0 % ",
    "0.0004 % ",
    "0.0014 %",
    "-",
    "-", 
    "-", 
    "-", 
    "-", 
    "-",  
    "-",  
    "-", 
    "-", 
    "-"
  ),
  missing_blank = c(
    "-", 
    "0.2 %",
    "0.2 %",
    "41.1 %",
    "8.1 %",
    "0.002 %", 
    "0.002 %", 
    "-", 
    "-",  
    "-",  
    "-", 
    "-", 
    "-"
  ),
  missing_zero = c(
    "-",
    "-",
    "-",
    "0.0006 %",
    "-", 
    "-", 
    "-", 
    "57.9 %", 
    "3.8 %",  
    "75.1 %",  
    "72.5 %", 
    "37.3 %",
    "-"
  ),
  complete = c(
    "72 %",
    "99.8 %",
    "99.8 %",
    "58.9 %", 
    "91.9 %", 
    "99.9 %", 
    "99.9 %", 
    "42.1 %",  
    "96.2 %",
    "24.9 %",  
    "27.5 %", 
    "62.7 %",
    "100 %" 
  )
)

names(payroll_data_missing_table) <- c("Column Name","NA","Blank","Zero", "Complete")

kable(payroll_data_missing_table,
      # caption = "Dataset Format", 
      align = "lcccc") |>
  kable_styling(full_width = FALSE) |>
  column_spec(1, width = "250px") |> 
  column_spec(2, width = "60px", background = ifelse(as.matrix(payroll_data_missing_table[, 2]) > 0, "#D7191C", ""),color = "white") |> 
  column_spec(3, width = "60px", background = ifelse(as.matrix(payroll_data_missing_table[, 3]) > 0, "#FDAE61", "")) |> 
  column_spec(4, width = "60px", background = ifelse(as.matrix(payroll_data_missing_table[, 4]) > 0, "#ABD9E9", "")) |> 
  column_spec(5, width = "60px", background = ifelse(as.matrix(payroll_data_missing_table[, 5]) > 0, "#2C7BB6", ""),color = "white") 
```


## Subset Details
```{r}
#| echo: false


payroll_data_missing_table <- tibble(
  column_name = c(
    "Mid Init",
    "Last Name, First Name",
    "Regular Hours", 
    "Regular Gross Paid", 
    "OT Hours", 
    "Total OT Paid", 
    "Total Other Pay",
    "Fiscal Year ,  Payroll Number , Agency Name, Agency Start Date, Work Location Borough, Title Description, Leave Status as of June 30, Base Salary, Pay Basis"
  ),
  
  
  missing_na = c(
    "-",
    "-",
    "-", 
    "-", 
    "-",  
    "-",  
    "-", 
    "-"
  ),
  missing_blank = c(
    "29.4 %",
    "1.2 %",
    "-", 
    "-", 
    "-",  
    "-",  
    "-", 
    "-"
  ),
  missing_zero = c(
    "-",
    "-",
    "6.9 %", 
    "1.5 %",  
    "20.9 %",  
    "15.9 %", 
    "3.7 %",
    "-"
  ),
  complete = c(
    "70.6 %",
    "98.8 %",
    "93.1 %", 
    "98.5 %",  
    "79.1 %",  
    "84.1 %", 
    "96.3 %",
    "100 %" 
  )
)

names(payroll_data_missing_table) <- c("Column Name","NA","Blank","Zero", "Complete")

kable(payroll_data_missing_table,
      # caption = "Dataset Format", 
      align = "lcccc") |>
  kable_styling(full_width = FALSE) |>
  column_spec(1, width = "250px") |> 
  column_spec(2, width = "60px", background = ifelse(as.matrix(payroll_data_missing_table[, 2]) > 0, "#D7191C", "")) |> 
  column_spec(3, width = "60px", background = ifelse(as.matrix(payroll_data_missing_table[, 3]) > 0, "#FDAE61", "")) |> 
  column_spec(4, width = "60px", background = ifelse(as.matrix(payroll_data_missing_table[, 4]) > 0, "#ABD9E9", "")) |> 
  column_spec(5, width = "60px", background = ifelse(as.matrix(payroll_data_missing_table[, 5]) > 0, "#2C7BB6", ""),color = "white") 
```


:::

Our original dataset has [■]{style="color:#D7191C;"} NA data for payroll_number, first_name, last_name. It also has [■]{style="color:#FDAE61;"} Blank data in 6 columns: first_name, mid_init, last_name, work_location_borough, title_description, agency_start_date.


For both datasets, [■]{style="color:#ABD9E9;"} Zero values show up in columns related to the amount paid to a person, which can be 3 categories: regular, overtime, other. If a certain category is 0, it means the person did not receive any money for that category. This in an important information because it does not requires us to exclude or impute these values. We can observe a significant reduction of zero values in the subset dataset compared to the original.

Our final subset data (Fire and Police Departments) no longer contains [■]{style="color:#D7191C;"} NA data, but still contains [■]{style="color:#FDAE61;"} Blank data in 3 columns: first_name, mid_init, last_name. 

For **mid_init** (Middle Name Initials), 29% of the data in the final subset is Blank, it is ok for us to have blanks in the data, this will not affect the future analysis. For **first_name** and **last_name** blank data (1.2% of the data in the final subset is Blank), these values are blank on purpose by the data provider: \
[*...As a part of FISA-OPA’s routine process for reviewing and releasing Citywide Payroll Data, data for some agencies (specifically NYC Police Department (NYPD) and the District Attorneys’ Offices (Manhattan, Kings, Queens, Richmond, Bronx, and Special Narcotics)) have been redacted since they are exempt from disclosure pursuant to the Freedom of Information Law, POL § 87(2)(f), on the ground that disclosure of the information could endanger the life and safety of the public servants listed thereon...*]{style="font-size:70%;"}

<!-- Based on this disclosure, we can confirm the Blank values for the names are part of the Police Department agency: -->
<!-- ```{r} -->
<!-- payroll_data_fire_police |>  -->
<!--   filter(first_name == "" | last_name == "") |>  -->
<!--   group_by(agency_name) |>  -->
<!--   summarise(`Count of Blank Records` = n()) |>  -->
<!--   kable(col.names = c("Agency", "Count of Blank Records"))  -->
<!-- ``` -->

\
In conclusion, the subset data for FIRE and POLICE departments (which will be used in the Results section) contains no [■]{style="color:#D7191C;"} NA values, only [■]{style="color:#FDAE61;"} Blank values that won't affect the analysis, [■]{style="color:#ABD9E9;"} Zero values that are crucial for salary-related insights, and multiple columns with a [■]{style="color:#2C7BB6;"} Complete set of values for all rows.



<!-- &nbsp;&nbsp;&nbsp; [■]{style="color:#D7191C;"}  NA      - any values equal to *NA* \ -->
<!-- &nbsp;&nbsp;&nbsp; [■]{style="color:#FDAE61;"}  Blank   - string values equal to "" \ -->
<!-- &nbsp;&nbsp;&nbsp; [■]{style="color:#ABD9E9;"}  Zero    - numeric values equal to 0 \ -->
<!-- &nbsp;&nbsp;&nbsp; [■]{style="color:#2C7BB6;"}  Complete - any values with data \ -->







\
\
\

<!-- ::: callout-note -->
<!-- **Instruction** -->

<!-- 2 Data data.qmd -->

<!-- 2.1 Technical description -->

<!-- Identify one or more data sources (see II. D. above) that you propose to draw on for the project. For each, describe how the data are collected and by whom. Describe the format of the data, the frequency of updates, dimensions, and any other relevant information. Note any issues / problems with the data, either known or that you discover. Explain how you plan to import the data. Carefully document your sources with links to the precise data sources that you used. If that is not possible (for example if your data is not available online, then explain that clearly.) -->

<!-- (suggested: 1/2 page) 250-500 words -->

<!-- 2.2 Missing value analysis -->

<!-- Describe any patterns you discover in missing values. If no values are missing, graphs should still be included showing that. -->

<!-- (suggested: 2 graphs plus commentary) -->
<!-- ::: -->

