# Data

For this project, we will be using a subset of the **Citywide Payroll Data**.\
This dataset is available via NYC OpenData and provided by the Office of Payroll Administration (OPA).

The final subset of data for this project will be based on 2 Agencies:\
<p style="margin-top: -15px">
-   Fire Department
-   Police Department
</p>

First, we will examine the metadata of the entire dataset, followed by an analysis of missing values to identify any potential issues that may impact our analysis. Finally, we will outline the decision-making process for selecting a subset of the data.

```{r}
#| echo: false

rm(list = ls())
```

```{r}
#| echo: false
#| message: false
#| warning: false

# Import Libraries
library(tibble)
library(knitr)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(visdat)
```


## Description

**Data Metadata**

```{r}
#| echo: false

# Create a tibble for the metadata
metadata <- tibble::tibble(
  Field = c(
    "Dataset Name",
    "Provided by",
    "Data Category",
    "Frequency of Updates", 
    "Date Created", 
    "Data Last Updated", 
    "Dimensions",
    "Each row represents",
    "Source URL"
  ),
  Value = c(
    "Citywide Payroll Data (Fiscal Year)",
    "Office of Payroll Administration (OPA)",
    "City Government",
    "Annually", 
    "October 31, 2015", 
    "October 30, 2024", 
    "6,225,611 rows and 17 columns",
    "City Employee Salary per Fiscal Year",
    "https://data.cityofnewyork.us/"
  )
)

# Print the metadata table using kable
kable(metadata, col.names = NULL, 
      # caption = "Dataset Metadata",
      align = "l")

```

\
\

**Data Format**

```{r}
#| echo: false



payroll_data_format <- tibble(
  column_name = c(
    "Fiscal Year", "Payroll Number", "Agency Name", "Last Name", "First Name", 
    "Mid Init", "Agency Start Date", "Work Location Borough", 
    "Title Description", "Leave Status as of June 30", "Base Salary", 
    "Pay Basis", "Regular Hours", "Regular Gross Paid", "OT Hours", 
    "Total OT Paid", "Total Other Pay"
  ),
  description = c(
    "Fiscal Year", "Payroll Number", "The Payroll agency that the employee works for", 
    "Last name of employee", "First name of employee", 
    "Middle initial of employee", "Date which employee began working for their current agency", 
    "Borough of employee's primary work location", "Civil service title description of the employee", 
    "Status of employee as of the close of the relevant fiscal year: Active, Ceased, or On Leave", 
    "Base Salary assigned to the employee", 
    "Lists whether the employee is paid on an hourly, per diem or annual basis", 
    "Number of regular hours employee worked in the fiscal year", 
    "The amount paid to the employee for base salary during the fiscal year", 
    "Overtime Hours worked by employee in the fiscal year", 
    "Total overtime pay paid to the employee in the fiscal year", 
    "Includes any compensation in addition to gross salary and overtime pay, i.e., Differentials, lump sums, uniform allowance, meal allowance, retroactive pay increases, settlement amounts, and bonus pay, if applicable."
  ),
  # api_field_name = c(
  #   "fiscal_year", "payroll_number", "agency_name", "last_name", "first_name", 
  #   "mid_init", "agency_start_date", "work_location_borough", 
  #   "title_description", "leave_status_as_of_june_30", "base_salary", 
  #   "pay_basis", "regular_hours", "regular_gross_paid", "ot_hours", 
  #   "total_ot_paid", "total_other_pay"
  # ),
  data_type = c(
    "Number", "Number", "Text", "Text", "Text", 
    "Text", "Timestamp", "Text", 
    "Text", "Text", "Number", 
    "Text", "Number", "Number", "Number", 
    "Number", "Number"
  )
)

names(payroll_data_format) <- c("Column Name","Description","Data Type")

kable(payroll_data_format,
      # caption = "Dataset Format", 
      align = "l") |>
  kable_styling(full_width = FALSE) |>
  column_spec(1, width = "200px")
```

\
\

**Importing the data**\

To access the data, we can go directly to the URL:\
<p style="font-size:95%; font-style:italic">
<https://data.cityofnewyork.us/City-Government/Citywide-Payroll-Data-Fiscal-Year-/k397-673e/about_data>
</p>

From this URL, you can click the button 'Export', and then select Download File-CSV format (All data 6225611 rows). This will download a file of about 840MB. Remember, we are using the data Last Updated on October 30, 2024.

The file will be downloaded with the name: Citywide_Payroll_Data\_\_Fiscal_Year\_\_**YYYYMMDD**.csv\
<p style="font-size:90%; font-style:italic">
(YYYYMMDD refers to the date that you downloaded the file)
</p>

Once the data was downloaded, it was added to a new folder inside the repository called 'data_source'.











```{r, cache=TRUE}
#| echo: false
#| message: false
#| warning: false
payroll_data_source <- read.csv("./data_source/Citywide_Payroll_Data__Fiscal_Year__20241111.csv") 
# payroll_data <- payroll_data_source
```



```{r}
#| echo: false
#| message: false
#| warning: false

payroll_data <- data.frame(payroll_data_source)

names(payroll_data) <- gsub("[\\.]+", "_", tolower(names(payroll_data_source)))

```






## Missing value analysis






```{r}
payroll_data_missing <- 
  payroll_data |>
  sample_n(100000) |>
  mutate(across(everything(), ~ if_else(. == "", NA, .)))
```

```{r}
payroll_data_missing |> 
  # select (-agency_name_clean, -work_location_borough_clean) |> 
  # sample_n(10000) |>
  vis_miss(
    sort_miss = TRUE,
    # cluster = TRUE,
    show_perc_col = TRUE,
    warn_large_data = FALSE,
    ) 
```


```{r}
payroll_data_total_summary <- payroll_data |>
  summarise(across(everything(), ~ n())) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Total Count"
  )

payroll_data_na_summary <- payroll_data |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "NA"
  )

payroll_data_blank_summary <- payroll_data |>
  summarise(across(everything(), ~ sum(. == "", na.rm = TRUE))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Blank"
  )

payroll_data_zero_summary <- payroll_data |>
  summarise(across(everything(), ~ sum(. == 0, na.rm = TRUE))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Zero"
  )
```




```{r}
payroll_data_combined_summary <- 
  payroll_data_total_summary |> 
  inner_join(payroll_data_na_summary, by = "Column") |> 
  inner_join(payroll_data_blank_summary, by = "Column") |> 
  inner_join(payroll_data_zero_summary, by = "Column")
```


```{r}
payroll_data_combined_summary
```


```{r}
payroll_data_combined_summary_likert <-
payroll_data_combined_summary |> 
  mutate(`Present` = `Total Count` - `NA` - `Blank` - Zero) |> 
  mutate(`Present` = `Present` / `Total Count`) |> 
  mutate(`NA` = `NA` / `Total Count`) |> 
  mutate(`Blank` = `Blank` / `Total Count`) |> 
  mutate(`Zero` = Zero / `Total Count`) |> 
  dplyr::select(`Column`,`NA`,`Blank`,Zero,`Present`) |> 
  filter(!Column %in% c('agency_name_clean','work_location_borough_clean'))

payroll_data_combined_summary_likert
```



```{r}
library(HH)
```


```{r}
#| fig-width: 10
#| fig-height: 5

likert_colors <- c("NA" = "#D7191C",
                   "Blank" = "#FDAE61",
                   "Zero" = "#ABD9E9",
                   "Present" = "#2C7BB6") 



HH::likert(x=Column~ .,payroll_data_combined_summary_likert, 
           positive.order=TRUE,
           as.percent = T,
           main = 'Diverging Stacked Bar',
           xlab='Percentage',
           ylab='Columns',
           col = likert_colors,
           ReferenceZero=2.5,
           rightAxis=FALSE
           )
```


```{r}
payroll_data_fire_police <- readRDS("data_source/payroll_data_fire_police.rds")

payroll_data_fire_police_total_summary <- payroll_data_fire_police |>
  summarise(across(everything(), ~ n())) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Total Count"
  )

payroll_data_fire_police_na_summary <- payroll_data_fire_police |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "NA"
  )

payroll_data_fire_police_blank_summary <- payroll_data_fire_police |>
  summarise(across(everything(), ~ sum(. == "", na.rm = TRUE))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Blank"
  )

payroll_data_fire_police_zero_summary <- payroll_data_fire_police |>
  summarise(across(everything(), ~ sum(. == 0, na.rm = TRUE))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Zero"
  )

```



```{r}
payroll_data_fire_police_combined_summary <- 
  payroll_data_fire_police_total_summary |> 
  inner_join(payroll_data_fire_police_na_summary, by = "Column") |> 
  inner_join(payroll_data_fire_police_blank_summary, by = "Column") |> 
  inner_join(payroll_data_fire_police_zero_summary, by = "Column")

```



```{r}
payroll_data_fire_police_combined_summary_likert <-
payroll_data_fire_police_combined_summary |> 
  mutate(`Present` = `Total Count` - `NA` - `Blank` - Zero) |> 
  mutate(`Present` = `Present` / `Total Count`) |> 
  mutate(`NA` = `NA` / `Total Count`) |> 
  mutate(`Blank` = `Blank` / `Total Count`) |> 
  mutate(`Zero` = Zero / `Total Count`) |> 
  dplyr::select(`Column`,`NA`,`Blank`,Zero,`Present`) |> 
  filter(!Column %in% c('agency_name_clean','work_location_borough_clean'))
```




```{r}
payroll_data_missing_group_likert <- bind_rows(
  payroll_data_combined_summary_likert |> mutate(group= 'Original Data'), 
  payroll_data_fire_police_combined_summary_likert |> mutate(group= 'Final Subset Data')
)

# payroll_data_missing_group_likert
```



```{r}
#| fig-width: 10
#| fig-height: 5

HH::likert(x=Column~. | group,payroll_data_missing_group_likert, 
           positive.order=TRUE,
           as.percent = T,
           main = 'Diverging Stacked Bar',
           xlab='Percentage',
           ylab='Columns',
           col = likert_colors,
           ReferenceZero=2.5,
           rightAxis=FALSE,
           xlimEqualLeftRight=TRUE
           )
```




```{r}
payroll_data_fire_police |> 
  filter(last_name == "") |> 
  group_by(agency_name) |> 
  summarise(`Count of Blank Records` = n()) |> 
  kable(col.names = c("Agency", "Count of Blank Records")) 
```






## Understanding and cleaning the raw data



```{r}
payroll_data |> 
  summarise(
    "Number of Agencies"     = n_distinct(agency_name),
    "Number of Titles"       = n_distinct(title_description),
    "Number of Locations"    = n_distinct(work_location_borough),
    "Number of Pay Basis"    = n_distinct(pay_basis),
    "Number of Fiscal Years" = n_distinct(fiscal_year)
  ) |> 
  kable()
```



### Agencies

Many of the agencies in the raw data have the following format: AGENCY_NAME #NUMBER.\
Examples:\
<p style="margin-top: -15px; font-size:85%;">
-   BROOKLYN COMMUNITY BOARD #1
-   BROOKLYN COMMUNITY BOARD #2
-   BROOKLYN COMMUNITY BOARD #3
</p>

We will group Agency names by aggregating all agencies that have the same name, but different numbers. This will reduce overall number os agencies.

```{r}
payroll_data <- 
  payroll_data |> 
  mutate(agency_name_clean = trimws(gsub("#\\d+$", "", agency_name)))

payroll_data |> 
  summarise(
    "Number of Agencies Before" = n_distinct(agency_name),
    "Number of Agencies After"  = n_distinct(agency_name_clean),
  ) |> 
  kable()
```

\
\

Now, let's take a look at the top 10 Agencies:

```{r}
payroll_data_summary <- payroll_data |> 
  group_by(agency_name_clean) |> 
  summarise(
    Total_records = n()
  ) |> 
  arrange(desc(Total_records)) |> 
  slice_head(n = 10)  


payroll_data_summary |> 
  kable(col.names = c("Agency", "Total Records")) |> 
  row_spec(which(payroll_data_summary$agency_name_clean == "FIRE DEPARTMENT"),   background = "#f94144", color = "white") |> 
  row_spec(which(payroll_data_summary$agency_name_clean == "POLICE DEPARTMENT"), background = "deepskyblue3", color = "white")


```

### Work Location Borough

When looking at the Work Location Borough, we see many recording with missing data (506,234 rows). We can also observe that the names of the locations are not standardized, some names are Uppercase and some are not (MANHATTAN vs Manhattan).

```{r}
payroll_data |> 
  group_by(work_location_borough) |> 
  summarise(
    Total_records = n()
  ) |> 
  arrange(desc(Total_records)) |> 
  slice_head(n = 10) |> 
  kable(col.names = c("Locations", "Total Records")) |> 
  row_spec(which(payroll_data |> 
                 group_by(work_location_borough) |> 
                 summarise(Total_records = n()) |> 
                 arrange(desc(Total_records)) |> 
                 slice_head(n = 10) |> 
                 pull(work_location_borough) %in% c("","Bronx", "MANHATTAN", "Manhattan")),   background = "#fcefb4", color = "black")
```


\
\

Our first step here is to capitalize all the location names.

```{r}
payroll_data <- payroll_data |> 
  mutate(work_location_borough_clean = toupper(work_location_borough))

```

```{r}
#| echo: false

# payroll_data |> 
#   group_by(work_location_borough_clean) |> 
#   summarise(
#     Total_records = n()
#   ) |> 
#   arrange((work_location_borough_clean)) |> 
#   slice_head(n = 10) |> 
#   kable(col.names = c("Locations", "Total Records"))
```

```{r}
payroll_data |> 
  summarise(
    "Number of Locations Before" = n_distinct(work_location_borough),
    "Number of Locations After"  = n_distinct(work_location_borough_clean),
  ) |> 
  kable()
```

\
\

Now let's take a look at the locations from the top 10 Agencies.



```{r}

required_boroughs <- c("BRONX", "BROOKLYN", "MANHATTAN", "QUEENS", "RICHMOND")

agencies_present_in_required_boroughs <-
  payroll_data |>
  filter(work_location_borough_clean %in% required_boroughs) |> 
  group_by(agency_name_clean) |>  
  summarise(
    borough_count = n_distinct(work_location_borough_clean)  
  ) |>
  filter(borough_count == length(required_boroughs)) |> 
  filter(agency_name_clean %in% payroll_data_summary$agency_name_clean) |> 
  pull(agency_name_clean)
```

```{r}

payroll_data_agencies_present_in_required_boroughs <-
  payroll_data |> 
  filter(agency_name_clean %in% payroll_data_summary$agency_name_clean) |> 
  filter(agency_name_clean %in% agencies_present_in_required_boroughs) |> 
  group_by(agency_name_clean) |> 
  summarise(
    Total_records = n(),
    "Number of Locations" = n_distinct(work_location_borough_clean)
  ) |> 
  arrange(desc(Total_records)) 

payroll_data_agencies_present_in_required_boroughs |> 
  kable(col.names = c("Agency", "Total Records", "Number of Locations")) |> 
  row_spec(which(payroll_data_agencies_present_in_required_boroughs$agency_name_clean == "FIRE DEPARTMENT"),   background = "#f94144", color = "white") |> 
  row_spec(which(payroll_data_agencies_present_in_required_boroughs$agency_name_clean == "POLICE DEPARTMENT"), background = "deepskyblue3", color = "white")

```

```{r}
payroll_data |> 
  filter(agency_name_clean %in%  c("POLICE DEPARTMENT","FIRE DEPARTMENT")) |> 
  filter(agency_name_clean %in% agencies_present_in_required_boroughs) |> 
  group_by(agency_name_clean,work_location_borough_clean,fiscal_year) |> 
  summarise(
    Total_records = n(), .groups = 'drop'
  ) |> 
  arrange(agency_name_clean,work_location_borough_clean,fiscal_year) |> 
  filter (!work_location_borough_clean %in% required_boroughs) |> 
  kable(col.names = c("Agency", "Location", "Fiscal Year", "Total Records"))
```

### Fiscal Year

```{r}
# Create the pivot table
pivot_table_fiscal_year <- payroll_data |>
filter(agency_name_clean %in%  c(
    "POLICE DEPARTMENT",
    "FIRE DEPARTMENT"
    )) |> 
  group_by(fiscal_year, agency_name_clean) |> 
  summarise(count = n(), .groups = 'drop') |> 
  pivot_wider(
    names_from = agency_name_clean,  
    values_from = count,  
    values_fill = list(count = 0) 
  )

pivot_table_fiscal_year |>
  kable(col.names = c("Fiscal Year", "FIRE DEPARTMENT", "POLICE DEPARTMENT")) |> 
  kable_styling() |>
  column_spec(3, background = ifelse(as.matrix(pivot_table_fiscal_year[, 3]) == 0, "#fcefb4", ""))
```

### Pay Basis

```{r}
payroll_data |> 
  # filter(fiscal_year == "2024") |>
  filter(agency_name_clean %in%  c(
    "POLICE DEPARTMENT",
    "FIRE DEPARTMENT"
    )) |> 
  group_by(agency_name_clean,pay_basis) |> 
  summarise(
    Total_records = n(),
    Count_of_titles = n_distinct(title_description)
    , .groups = 'drop'
  ) |> 
  arrange(desc(Total_records)) |>
  kable(col.names = c("Agency", "Pay Basis", "Total Records", "Count of Titles"))
```

```{r}
payroll_data |> 
  # filter(fiscal_year == "2024") |>
  filter(agency_name_clean %in%  c(
    "POLICE DEPARTMENT",
    "FIRE DEPARTMENT"
    )) |> 
  filter(pay_basis == "Prorated Annual") |>
  filter(title_description == "CHAPLAIN") |> 
  group_by(agency_name_clean,pay_basis,title_description) |> 
  summarise(
    Total_records = n()
    , .groups = 'drop'
  ) |> 
  arrange(agency_name_clean) |>
  kable(col.names = c("Agency", "Pay Basis", "Title", "Total Records"))
```


```{r}
payroll_data_filtered <- 
  payroll_data |> 
  filter(agency_name_clean %in%  c("POLICE DEPARTMENT","FIRE DEPARTMENT")) |> 
  filter(work_location_borough_clean %in%   c("BRONX", "BROOKLYN", "MANHATTAN", "QUEENS", "RICHMOND")) |> 
  filter(fiscal_year > "2014") |> 
  filter(pay_basis %in%   c("per Annum", "per Day", "per Hour"))
  
```



```{r}

# Count NA values and pivot
na_summary <- payroll_data_filtered |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "NA Count"
  )

# Count blank values and pivot
blank_summary <- payroll_data_filtered |>
  summarise(across(everything(), ~ sum(. == "", na.rm = TRUE))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Blank Count"
  )

# Count zero values and pivot
zero_summary <- payroll_data_filtered |>
  summarise(across(everything(), ~ sum(. == 0, na.rm = TRUE))) |>
  pivot_longer(
    cols = everything(),
    names_to = "Column",
    values_to = "Zero Count"
  )

# Combine all three summaries into one dataframe
combined_summary <- na_summary |>
  inner_join(blank_summary, by = "Column") |>
  inner_join(zero_summary, by = "Column")

# Display the combined result using kable
combined_summary |>
  filter(`NA Count` > 0 | `Blank Count` > 0 | `Zero Count` > 0) |> 
  kable(col.names = c("Column", "NA Count", "Blank Count", "Zero Count"))

```

### Payroll Number	

```{r}
payroll_data_filtered_payroll <- payroll_data_filtered |> 
  group_by(agency_name_clean,payroll_number,fiscal_year) |> 
  summarise(
    Total_records = n()
    ,.groups = "drop"
  ) |> 
  arrange(agency_name_clean,fiscal_year,desc(Total_records))

payroll_data_filtered_payroll |> 
  kable(col.names = c("Agency", "Payroll Number", "Fiscal Year", "Total Records")) |>
  column_spec(2, background = ifelse(is.na(as.matrix(payroll_data_filtered_payroll[, 2])), "#fcefb4", ""))
```



```{r}
payroll_data_fire_police <- 
  payroll_data_filtered |>
  mutate(
    payroll_number = case_when(
      is.na(payroll_number) & agency_name_clean == "FIRE DEPARTMENT" ~ 57,
      is.na(payroll_number) & agency_name_clean == "POLICE DEPARTMENT" ~ 56,
      TRUE ~ payroll_number
    )
  ) |> 
  mutate(
    agency_name = agency_name_clean,
    work_location_borough = work_location_borough_clean
  ) |>
  dplyr::select(-agency_name_clean, -work_location_borough_clean)
```


### Agency Start Date

```{r}
library(lubridate)

payroll_data_fire_police <- 
  payroll_data_fire_police |> 
  mutate(agency_start_date = mdy(agency_start_date))
```





**Save final Data:**\

```{r}
# CSV
# write.csv(payroll_data_fire_police, "data_source/payroll_data_fire_police.csv", row.names = FALSE)

# RDS
# saveRDS(payroll_data_fire_police, "data_source/payroll_data_fire_police.rds")
# payroll_data_fire_police <- readRDS("data_source/payroll_data_fire_police.rds")
```


Missing value stuff goes here.

## Understanding and cleaning the raw data

Note any issues / problems with the data, either known or that you discover.

## Deciding on final data subset

Based on the analysis above, we will add a few filter to reduce and clean the dataset.

**Filters applied to the dataset:**\

```{r}
#| echo: false

library(tibble)
library(knitr)
library(kableExtra)

payroll_data_format <- tibble(
  column_name = c(
    "Fiscal Year", 
    "Agency Name", 
    "Pay Basis"
  ),
  filter = c(
    "From 2015 to 2024", 
    "FIRE DEPARTMENT, POLICE DEPARTMENT",
    "per Annum, per Day, per Hour"
  )
)

names(payroll_data_format) <- c("Column Name","Filter")

kable(payroll_data_format,
      # caption = "Dataset Format", 
      align = "l") |>
  kable_styling(full_width = FALSE) |>
  column_spec(1, width = "200px")
```

\
\

**Cleaning data and Handling NAs:**\

```{r}
#| echo: false

library(tibble)
library(knitr)
library(kableExtra)

payroll_data_format <- tibble(
  column_name = c(
    "Fiscal Year", 
    "Agency Name",
    "Payroll Number",
    "Work Location Borough", 
    "Pay Basis"
  ),
  filter = c(
    "2014 had missing data for Police Department. 2014 will be removed.", 
    "Group agencies with similar names.",
    "Missing values have been backfilled based on Agency Name.",
    "Uppercase all names. Filtering data only for top 5 Boroughs.",
    "Keeping only the top 3 Pay Basis."
  )
)

names(payroll_data_format) <- c("Column Name","Note")

kable(payroll_data_format,
      # caption = "Dataset Format", 
      align = "l") |>
  kable_styling(full_width = FALSE) |>
  column_spec(1, width = "200px")
```

\
\

**Size of Final subset Data**\

```{r}
#| echo: false

library(knitr)
library(kableExtra)

# Create a tibble for the metadata
metadata <- tibble::tibble(
  Field = c(
    # "Dataset Name",
    # "Provided by",
    # "Data Category",
    # "Frequency of Updates", 
    # "Date Created", 
    # "Data Last Updated", 
    "Dimensions",
    "Each row represents"
    # "Source URL"
  ),
  Value = c(
    # "Citywide Payroll Data (Fiscal Year)",
    # "Office of Payroll Administration (OPA)",
    # "City Government",
    # "Annually", 
    # "October 31, 2015", 
    # "October 30, 2024", 
    "804,630 rows and 17 columns",
    "City Employee Salary per Fiscal Year"
    # "https://data.cityofnewyork.us/"
  )
)

# Print the metadata table using kable
kable(metadata, col.names = NULL, 
      # caption = "Dataset Metadata",
      align = "l")

```

\

**Record count by Agency:**\

```{r}
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 3

library(ggplot2)
library(tibble)
library(dplyr)

payroll_data_fire_police |> 
  group_by(agency_name) |> 
  summarise(
    Count = n()
    , .groups = 'drop'
  ) |> 
  arrange(desc(Count)) |>
  mutate(Percentage = Count / sum(Count) * 100,
         Label = paste(
                       format(Count, big.mark = ","), "\n", " (", sprintf("%.0f", Percentage), "%)", sep = "")) |> 
ggplot(aes(x = "", y = Count, fill = agency_name)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  scale_fill_manual(name= 'Agency', values = c("#f94144", "deepskyblue3")) +
  # labs(title = "Record Count by Department") +
  theme_void() +
  geom_text(aes(label = Label), position = position_stack(vjust = 0.5), color = "white")

```


**Overall dataset Information:**\

```{r}
#| echo: false


# Create a tibble for the metadata
metadata_final_table <- tibble::tibble(
  "Metrics" = c(
    "Count of Records",
    "Fiscal Years",
    "Agency Start Date Range", 
    "Work Location Borough",
    "Title Description"
  ),
  "Fire Deparmtnet" = c(
    "192,638",
    "10 years (2015-2024)",
    "1968-04-22 to 2024-07-29   ", 
    "5",
    "238"
  ),
  "Police Department" = c(
    "611,992",
    "10 years (2015-2024)",
    "1960-11-14 to 2049-10-16", 
    "5",
    "335"
  ),
)



# Print the metadata table using kable
kable(metadata_final_table, 
      # col.names = NULL, 
      # caption = "Dataset Metadata",
      align = "l")

```

\
\
\
\
\
\
\
\
\

::: callout-note
**Instruction**

2 Data data.qmd

2.1 Technical description

Identify one or more data sources (see II. D. above) that you propose to draw on for the project. For each, describe how the data are collected and by whom. Describe the format of the data, the frequency of updates, dimensions, and any other relevant information. Note any issues / problems with the data, either known or that you discover. Explain how you plan to import the data. Carefully document your sources with links to the precise data sources that you used. If that is not possible (for example if your data is not available online, then explain that clearly.)

(suggested: 1/2 page) 250-500 words

2.2 Missing value analysis

Describe any patterns you discover in missing values. If no values are missing, graphs should still be included showing that.

(suggested: 2 graphs plus commentary)
:::
